---
title: "lab01-birthwt-homework.knit"
output: html_document
---

# 자료읽기

```{r}
library(tidyverse)
library(caret)
DF <- read.csv('C:\\WORK\\data\\birthwt4times.csv')
DF <-
DF %>%
mutate(
 low = factor(low),
 lwtkg = round(lwt*0.453592,1),
 race = factor(race))
DF$lwt <- NULL
```

# 자료분할

```{r}
TR <- DF[seq(1, nrow(DF), 2),]
dim(TR)

TS <- DF[seq(2, nrow(DF), 2),]
dim(TS)
```

# 1. 선형회귀 모형

## age vs bwt
```{r}
ggplot(TR, aes(x=age, y=bwt)) +
geom_point() +
 geom_hline(yintercept=2500) +
 geom_smooth(method='lm')
```

## lwtkg vs bwt
```{r}
ggplot(TR, aes(x=lwtkg, y=bwt)) +
geom_point() +
 geom_hline(yintercept=2500) +
 geom_smooth(method='lm')
```

## lwtkg vs bwt | race
```{r}
ggplot(TR, aes(x=lwtkg, y=bwt)) +
geom_point() +
 geom_smooth(method='lm') +
 facet_wrap(~race)
```

## race vs bwt
```{r}
ggplot(TR, aes(x=race, y=bwt)) +
 geom_boxplot()
```

## smoke vs bwt
```{r}
ggplot(TR, aes(x=factor(smoke), y=bwt)) +
 geom_boxplot()
```

## smoke vs bwt | smoke
```{r}
ggplot(TR, aes(x=age, y=bwt)) +
geom_point() +
 geom_smooth(method='lm') +
 facet_wrap(~smoke)
```

## ui vs bwt
```{r}
ggplot(TR, aes(x=factor(ui), y=bwt)) +
 geom_boxplot()
```

## ht vs bwt
```{r}
ggplot(TR, aes(x=factor(ht), y=bwt)) +
 geom_boxplot()
```

# 모형적합
- 종속변수(y): bwt
- 독립변수: age, ftv, ptl, race, smoke, ht, ui, lwtkg

```{r}
Rlm <- lm(bwt ~ age+ftv+ptl+race+smoke+ht+ui+lwtkg, data=TR)
summary(Rlm)
```

```{r}
TROUT <-
 TR %>%
 mutate(
 yh=predict(Rlm),
 e=residuals(Rlm))
head(TROUT)
```

```{r}
mean(TROUT$e^2) # MSE
```

```{r}
mean(abs(TROUT$e)) # MAE
```

# 모형검토(TR)

```{r}
ggplot(TROUT, aes(x=bwt, y=yh)) +
geom_point() +
geom_smooth(method='lm')
```

```{r}
ggplot(TROUT, aes(x=yh, y=e)) +
geom_point() +
 geom_smooth(method='lm')
```

```{r}
ggplot(TROUT, aes(x=e)) +
geom_histogram()
```

```{r}
TROUT %>% summarize(mn=mean(e), sd=sd(e), min=min(e), max=max(e))
```

```{r}
summary(TROUT$e)
```

# 모형평가(TS)

```{r}
TSOUT <-
TS %>%
mutate(yh=predict(Rlm, TS), e=bwt-yh)
head(TSOUT)
```

```{r}
mean(TSOUT$e^2) # MSE
```

```{r}
mean(abs(TSOUT$e)) # MAE
```

```{r}
TSOUT %>% summarize(mn=mean(e), sd=sd(e), min=min(e), max=max(e))
```

```{r}
summary(TSOUT$e)
```

# 스코어

```{r}
SC <- read.csv(text='
age,ftv,race,ptl,smoke,ht,ui,lwtkg
30,0,3,0,1,0,0,60
40,0,3,0,1,0,0,60
30,0,3,0,0,0,0,60
40,0,3,0,0,0,0,60
30,0,3,0,1,1,0,60
40,0,3,0,1,1,0,60
30,0,3,0,0,1,0,60
40,0,3,0,0,1,0,60
30,0,3,0,1,1,1,60
40,0,3,0,1,1,1,60
30,0,3,0,0,1,1,60
40,0,3,0,0,1,1,60
')
SC <- SC %>% mutate(race=factor(race, levels=1:3))
SCOUT <-
 SC %>%
 mutate(yh = predict(Rlm, SC))
SCOUT
```

# 2. 로지스틱회귀모형
## age vs low
```{r}
ggplot(TR, aes(x=factor(low), y=age)) +
geom_boxplot(fill='gray') +
 geom_jitter(color='red', alpha=0.5, size=2)
```

## lwtkg vs low
```{r}
ggplot(TR, aes(x=factor(low), y=lwtkg)) +
geom_boxplot(fill='gray') +
 geom_jitter(color='red', alpha=0.5, size=2)
```

## race vs low
```{r}
ggplot(TR, aes(x=race, fill=low)) +
geom_bar()
```

## smoke vs low
```{r}
ggplot(TR, aes(x=factor(smoke), fill=low)) +
geom_bar()
```


# 모형적합
- 종속변수(y): low
- 독립변수: age, ftv, ptl, race, smoke, ht, ui, lwtkg

```{r}
Mglm <- glm(low ~ age+ftv+ptl+race+smoke+ht+ui+lwtkg, data=TR, family=binomial)
summary(Mglm)
```

# 모형검토(TR)

```{r}
TROUT <-
 TR %>% dplyr::select(low) %>%
 mutate(
 ph = predict(Mglm, type='response'),
 yh = factor(ifelse(ph>=0.5, 1, 0)))
head(TROUT)
```

```{r}
confusionMatrix(TROUT$yh, TROUT$low, positive='1')
```

# 모형평가(TS)

```{r}
TSOUT <-
TS %>%
mutate(yh=predict(Rlm, TS), e=bwt-yh)
head(TSOUT)
```

```{r}
mean(TSOUT$e^2) # MSE
```

```{r}
mean(abs(TSOUT$e)) # MAE
```

```{r}
TSOUT %>% summarize(mn=mean(e), sd=sd(e), min=min(e), max=max(e))
```

```{r}
summary(TSOUT$e)
```


```{r}
TSOUT <-
 TS %>% dplyr::select(low) %>%
 mutate(
 ph = predict(Mglm, TS, type='response'),
 yh = factor(ifelse(ph>=0.5, 1, 0)))
head(TSOUT)
```

```{r}
confusionMatrix(TSOUT$yh, TSOUT$low, positive='1')
```

# 스코어

```{r}
SC <- read.csv(text='
age,ftv,race,ptl,smoke,ht,ui,lwtkg
30,0,3,0,1,0,0,60
40,0,3,0,1,0,0,60
30,0,3,0,0,0,0,60
40,0,3,0,0,0,0,60
30,0,3,0,1,1,0,60
40,0,3,0,1,1,0,60
30,0,3,0,0,1,0,60
40,0,3,0,0,1,0,60
30,0,3,0,1,1,1,60
40,0,3,0,1,1,1,60
30,0,3,0,0,1,1,60
40,0,3,0,0,1,1,60
')
SC <- SC %>% mutate(race=factor(race, levels=1:3))
SCOUT <-
 SC %>%
 mutate(
 ph = predict(Mglm, SC, type='response'),
 yh = factor(ifelse(ph>=0.5, 1, 0)))
SCOUT
```

# 3. 나무모형

```{r}
library(yardstick)
library(ROCR)
library(pROC)
library(rpart)
library(rpart.plot)
```

```{r}
TR %>%
group_by(low) %>%
summarize_if(is.numeric, 
             list(mn='mean', sd='sd', min='min', max='max'))
```

# 모형적합
- 종속변수(y): low
- 독립변수: age, ftv, ptl, race, smoke, ht, ui, lwtkg

```{r}
Mr <- rpart(low~ age+ftv+ptl+race+smoke+ht+ui+lwtkg, data=TR)
summary(Mr)
```

```{r}
plot(Mr, margin=1)
text(Mr, use.n=TRUE)
rpart.plot(Mr, type=4)
```

# 모형검토(TR)

```{r}
TROUT <-
 TR %>% dplyr::select(low) %>%
 mutate(
 ph = predict(Mr, type='prob')[,2],
 yh = factor(ifelse(ph>=0.5, 1, 0)))
head(TROUT)
```

```{r}
confusionMatrix(TROUT$yh, TROUT$low, positive='1')
```

```{r}
TR <-
TR %>%
  mutate(
    ph = predict (Mr, type='prob')[,2],
    yh = factor(ifelse(ph>=0.5, 1, 0)))
head(TR)
```

# 모형평가(TS)

```{r}
TSOUT <-
TS %>%
mutate(yh=predict(Mr, TS), e=bwt-yh)
head(TSOUT)
```

```{r}
mean(TSOUT$e^2) # MSE
```

```{r}
mean(abs(TSOUT$e)) # MAE
```

```{r}
TSOUT %>% summarize(mn=mean(e), sd=sd(e), min=min(e), max=max(e))
```

```{r}
summary(TSOUT$e)
```

# 스코어

```{r}
SC <- read.csv(text='
age,ftv,race,ptl,smoke,ht,ui,lwtkg
30,0,3,0,1,0,0,60
40,0,3,0,1,0,0,60
30,0,3,0,0,0,0,60
40,0,3,0,0,0,0,60
30,0,3,0,1,1,0,60
40,0,3,0,1,1,0,60
30,0,3,0,0,1,0,60
40,0,3,0,0,1,0,60
30,0,3,0,1,1,1,60
40,0,3,0,1,1,1,60
30,0,3,0,0,1,1,60
40,0,3,0,0,1,1,60
')
SC <- SC %>% mutate(race=factor(race, levels=1:3))

SCOUT <-
  SC %>%
  mutate(
    ph = predict(Mr, SC, type='prob')[,2],
    yh = factor(ifelse(ph>=0.5, 1, 0)))
SCOUT
```


# 4. 랜덤포레스트

```{r}
library(randomForest)
```

# 모형적합
- 종속변수(y): low
- 독립변수: age, ftv, ptl, race, smoke, ht, ui, lwtkg

```{r}
Mrf <- randomForest(low~ age+ftv+ptl+race+smoke+ht+ui+lwtkg, data=TR)
summary(Mr)
```

```{r}
plot(Mrf)
varImpPlot(Mrf)
treesize(Mrf)
```

# 모형검토(TR)

```{r}
TROUT <-
 TR %>% dplyr::select(low) %>%
 mutate(
 ph = predict(Mrf, type='prob')[,2],
 yh = factor(ifelse(ph>=0.5, 1, 0)))
head(TROUT)
```

```{r}
confusionMatrix(TROUT$yh, TROUT$low, positive='1')
```

```{r}
TR <-
TR %>%
  mutate(
    ph = predict (Mrf, type='prob')[,2],
    yh = factor(ifelse(ph>=0.5, 1, 0)))
head(TR)
```

# 모형평가(TS)

```{r}
TSOUT <-
TS %>%
mutate(yh=predict(Mrf, TS), e=bwt-yh)
head(TSOUT)
```

```{r}
mean(TSOUT$e^2) # MSE
```

```{r}
mean(abs(TSOUT$e)) # MAE
```

```{r}
TSOUT %>% summarize(mn=mean(e), sd=sd(e), min=min(e), max=max(e))
```

```{r}
summary(TSOUT$e)
```

# 스코어

```{r}
SC <- read.csv(text='
age,ftv,race,ptl,smoke,ht,ui,lwtkg
30,0,3,0,1,0,0,60
40,0,3,0,1,0,0,60
30,0,3,0,0,0,0,60
40,0,3,0,0,0,0,60
30,0,3,0,1,1,0,60
40,0,3,0,1,1,0,60
30,0,3,0,0,1,0,60
40,0,3,0,0,1,0,60
30,0,3,0,1,1,1,60
40,0,3,0,1,1,1,60
30,0,3,0,0,1,1,60
40,0,3,0,0,1,1,60
')
SC <- SC %>% mutate(race=factor(race, levels=1:3))

SCOUT <-
  SC %>%
  mutate(
    ph = predict(Mrf, SC, type='prob')[,2],
    yh = factor(ifelse(ph>=0.5, 1, 0)))
SCOUT
```
